### 2013-08-26 doc/dissertation/document/chapters/conclusion/conclusion.tex
adding pauls corrections to the conclusion

It was found that the leakage of histories from domain to domain in
the parallel Monte Carlo algorithm could in fact be quantified
analytically using the algebraic properties of the system. These
relationships were then used to determine the amount of overlap one
may require in the parallel algorithm to reduce communication costs
and increase parallel efficiencies. Scaling studies showed that in the
strong case overlap in small quantities on the order of the
mean-free-path of a stochastic history in the simulation could boost
parallel efficiencies by a few percent. However, it was found that
this additional overlap was not effective in boosting strong scaling
efficiencies. It was found that in general overlap was not very
effective due to the fact that the parallel communication saved during
the Monte Carlo transport sequence is simply deferred until after
transport is complete when it manifests itself as an overlapping
parallel vector reduction operation.

@@

Finally, given that overlap was not very constructive in boosting
parallel efficiencies, it was postulated that very little
domain-to-domain communication of histories was occurring in the first
place. Using this idea, we implemented a subdomain Neumann-Ulam method
with MCSA such that MCSA now takes the form of a stochastic
realization of an additive Schwarz method. Scaling studies using this
technique showed significant improvements in parallel scalability even
when compared to the MCSA results when domain-to-domain communication
was present, further enhancing the method and meeting the goal of
improved scalability.

@@

These operations were observed to substantially increase runtimes of
MCSA and consume unacceptable amounts of memory, mostly due to the
fact that inverting the preconditioners created dense
matrices. Although a good fraction of the elements in each row of the
composite operator could be eliminated through the reduced domain
approximation developed by this work, memory costs still prevented
higher fidelity simulations from being performed. In addition, the
reduced domain approximation does not prevent the large run times
incurred for inverting the preconditioners. We therefore propose the
following two developments in preconditioning for Monte Carlo
methods.

\subsubsection{Variance Reduction-Based Preconditioning}
Traditional Monte Carlo variance reduction techniques instead of
preconditioning strategies that rely on algebraic preconditioners
should be investigated. Constructing the complete inverse of the
preconditioner is the core time and scalability constraint when modern
algebraic preconditioners are used. Building the inverse for
preconditioners where a sparsity pattern is not enforced typically
results in a dense matrix that destroys scalability for domain
decomposed parallel computations. Alleviating both of these issues by
stochastic preconditioning methods should be studied. Decades of
research and development in Monte Carlo variance reduction for
particle transport \cite{booth_1994} may be leveraged to form similar
strategies for linear systems. In the same way that this work has been
able to adapt modern particle transport strategies for parallelism and
apply them to linear systems, the same strategy may possibly be
applied to variance reduction techniques to improve MCSA convergence
by modifying the weights and probabilities of the Monte Carlo game as
an effective means of preconditioning.

@@

\subsection{Breaking Away from $\rho(H) < 1$}
\label{subsec:future_spec_rad}
All numerical methods for solving linear problems require quality
preconditioning. Even a Krylov method may be effectively useful
without it. However, these methods typically converge eventually
whereas quality preconditioning may not be enough to guarantee
convergence of MCSA. We saw in our work that in many cases very
aggressive preconditioning was not enough to condition the eigenvalues
of the system into a regime where MCSA would apply. The limitation
that the spectral radius of the physics operator iteration matrix must
be less than one is by far the most prohibitive component of MCSA. The
fact that the Monte Carlo sequence is based on the Richardson
iteration means that without sufficient preconditioning, even if this
condition is achieved convergence may be incredibly slow. Therefore,
we propose future work that aims to move away from this restriction
that is fundamentally based on sampling a sequence not based on the
Neumann series.

>>>>

Outlined in Appendix~\ref{chap:parallel_theory}, it was found that the
leakage of histories from domain to domain in the parallel Monte Carlo
algorithm could in fact be quantified analytically using the algebraic
properties of the system. These relationships were then used to
determine the amount of overlap one may require in the parallel
algorithm to reduce communication costs and increase parallel
efficiencies. Scaling studies showed that in the strong case overlap
in small quantities on the order of the mean-free-path of a stochastic
history in the simulation could boost parallel efficiencies by up to
10\% in isolated cases. However, it was found that this additional
overlap was not effective in boosting weak scaling efficiencies. In
general, overlap was not very effective due to the fact that the
parallel communication saved during the Monte Carlo transport sequence
is simply deferred until after transport is complete when it manifests
itself as an overlapping parallel vector reduction operation. As
compared to transport calculations where this overlap procedure was
very effective, in the context of MCSA the Monte Carlo calculations
are significantly shorter with $O(10,000)$ histories used in this
work. These shorter calculations and more frequent overlapping tally
vector reductions create an overhead that is not observed in the
literature for transport calculations.

@@

Finally, given that overlap was not very constructive in boosting
parallel efficiencies, it was postulated that very little
domain-to-domain communication of histories was occurring in the first
place. Motivated by this idea, we implemented a subdomain Neumann-Ulam
method with MCSA such that MCSA now takes the form of a stochastic
realization of an additive Schwarz method. Scaling studies using this
technique showed significant improvements in parallel scalability even
when compared to the MCSA results when domain-to-domain communication
was present, further enhancing the method and meeting the goal of
improved scalability.

@@

These operations were observed to substantially increase run-times of
MCSA and consume unacceptable amounts of memory, mostly due to the
fact that inverting the preconditioners created dense
matrices. Although a good fraction of the elements in each row of the
composite operator could be eliminated through the reduced domain
approximation developed by this work, memory costs still prevented
higher fidelity simulations from being performed. In addition, the
reduced domain approximation does not prevent the large run times
incurred for inverting the preconditioners. We therefore propose the
following two developments in preconditioning for Monte Carlo
methods.

\subsubsection{Variance Reduction-Based Preconditioning}
Traditional Monte Carlo variance reduction techniques instead of
preconditioning strategies that rely on algebraic preconditioners
should be investigated. Constructing the complete inverse of the
preconditioner is the core time and scalability constraint when modern
algebraic preconditioners are used. Building the inverse for
preconditioners where a sparsity pattern is not enforced typically
results in a dense matrix that destroys scalability for domain
decomposed parallel computations. Alleviating both of these issues by
stochastic preconditioning methods should be studied. Decades of
research and development in Monte Carlo variance reduction for
particle transport \cite{booth_1994} may be leveraged to form similar
strategies for linear systems. In the same way that this work has been
able to adapt modern particle transport strategies for parallelism and
apply them to linear systems, the same strategy may possibly be
applied to variance reduction techniques to improve MCSA convergence
by modifying the weights and probabilities of the Monte Carlo game,
effectively serving as a form of preconditioning by modifying the
Neumann-Ulam decomposition and subsequently the physics operator.

@@

\subsection{Breaking Away from $\rho(H) < 1$}
\label{subsec:future_spec_rad}
All numerical methods for solving linear problems require quality
preconditioning. Even a Krylov method may be effectively useless
without it. However, these methods typically converge eventually
whereas quality preconditioning may not be enough to guarantee
convergence of MCSA. We saw in our work that in many cases very
aggressive preconditioning was not enough to condition the eigenvalues
of the system into a regime where MCSA would apply. The limitation
that the spectral radius of the physics operator iteration matrix must
be less than one is by far the most prohibitive component of MCSA. The
fact that the Monte Carlo sequence is based on the Richardson
iteration means that without sufficient preconditioning, even if this
condition is achieved convergence may be incredibly slow. Therefore,
we propose future work that aims to move away from this restriction
that is fundamentally based on sampling a sequence not based on the
Neumann series.
